---
title: "data_exploration"
subtitle: "My team's name goes here"
author: "The team members' names go here"
date: "`r Sys.time()`"
urlcolor: blue # to show hyperlinks in blue when printed as pdf

# uncomment below to render to html
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
# uncomment below to render to pdf
# output:
#   bookdown::pdf_book:
#     toc: true
#     toc_depth: 4
#     highlight: tango
    
bibliography: [references/packages.bib, references/references.bib]
biblio-style: apalike
nocite: '@*'
---

```{r setup, echo=FALSE, message=FALSE}
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse
library("cjoint")
library("marginaleffects")
library("emmeans")
library("broom")
library("brms")
library("ggdist")
library("ggforce")
library("scales")

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

# include references for used packages
write_bib(.packages(), "references/packages.bib") 
```

# Moral Machines Experimental Paradigm

# Exploration of methodology used in original study

```{r message=F}
# Load up common functions from original paper
source("MMFunctionsShared.R")
```

## AMCE

```{r}
# Original Paper's analysis of the main effects
# Take a small sample of ~100K paired rows instead of
# working with the full dataset due to size & memory constraints
df.small_profiles = fread(input="../../data/SharedResponses_small.csv")

profiles <- PreprocessProfiles(df.small_profiles)
```

```{r message=F, warning=F}
# Main Figure 2 - (a)
Coeffs.main <- GetMainEffectSizes(profiles,T,9)
plotdata.main <- GetPlotData(Coeffs.main,T,9)

Coeffs.util <- GetMainEffectSizes.Util(profiles)
plotdata.util <- GetPlotData.Util(Coeffs.util)

## Save them
save(plotdata.main, file = "plotdatamain.rdata")
save(plotdata.util, file = "plotdatautil.rdata")
## Plot them
plot <- PlotAndSave(plotdata.main,T,"MainChangePr",plotdata.util)

# Show plot
plot
```

### Own-self implementation of the analysis used (sanity-check)

```{r}
lev = levels(profiles$ScenarioType)
print(lev)

AttLevels = levels(profiles$AttributeLevel)
print(AttLevels)

# Focusing on the Age characteristic
df.age_profiles = profiles %>%
  filter(ScenarioType==lev[1] & profiles$ScenarioTypeStrict==lev[1]) %>% 
  mutate(AttributeLevel = factor(AttributeLevel, levels=AttLevels[(4*2):(4*2+1)]))

df.age_profiles %>% head()

fit.age = lm(Saved ~ as.factor(AttributeLevel), data=df.age_profiles)

fit.age %>% tidy()
```

This is almost matching, with the original study method producing 0.5097648 while our naive lm method produces 0.5312953. We're still missing the 'weighting' part of this method. Let's take a peek under the hood for this `calcWeightsTheoretical` function:

```{r eval=F}
# This returns the inverse probability (1/p) of the data distribution
calcWeightsTheoretical <- function(profiles){
  p <- apply(profiles,1,CalcTheoreticalInt)
  return(1/p)
}

# A-priori random generation probability constraints on some of the
# other attributes (Intervention, Barrier, PedPed, CrossingSignal).
# Which relates to the interaction between legality of crossing and
# whether there are passengers in the car.
CalcTheoreticalInt <- function(X){
  if (X["Intervention"]==0){
    if (X["Barrier"]==0){
      if (X["PedPed"] == 1) p <- 0.48
      else p <- 0.32
      
      if (X["CrossingSignal"]==0) p <- p*0.48
      else if (X["CrossingSignal"]==1) p <- p*0.2
      else p <- p * 0.32
    }
    else p <- 0.2
  }
  else {
    if (X["Barrier"]==0){
      if (X["PedPed"] == 1) {
        p <- 0.48
        if (X["CrossingSignal"]==0) p <- p*0.48
        else if (X["CrossingSignal"]==1) p <- p*0.32
        else p <- p * 0.2
      }
      else {
        p <- 0.2
        if (X["CrossingSignal"]==0) p <- p*0.48
        else if (X["CrossingSignal"]==1) p <- p*0.2
        else p <- p * 0.32
      }
    }
    else p <- 0.32
  }
  return(p)
}
```

And incorporate these weights into our lm model:

```{r}
fit.age = lm(Saved ~ as.factor(AttributeLevel),
             data=df.age_profiles,
             weights=calcWeightsTheoretical(df.age_profiles))

fit.age %>% tidy()
```

We recover exactly: 0.5097648 which is similarly the value given by the study's code implementation.

## Brief note on glm logistic regression

Now, what if we were to perform a logistic regression instead of a lm? Since the outcome is limited to 0 or 1. However, this would be uninteresting as our independent variables are categorical and not continuous.

```{r}
fit.logit_age = glm(Saved ~ as.factor(AttributeLevel),
                    data=df.age_profiles,
                    weights=calcWeightsTheoretical(df.age_profiles))

fit.logit_age %>% tidy()
```

No difference on accuracy nor error statistics from the lm model.

## brms bayesian analysis

```{r}
# Re-code CrossingSignal to be -1, 0, 1 with 0 as no legal
# -1 as illegal
# 1 as legal
profiles$CrossingSignal <- as.numeric(levels(profiles$CrossingSignal))[profiles$CrossingSignal]
profiles$CrossingSignal[profiles$CrossingSignal == 2] <- -1

profiles <- profiles %>% 
  mutate(CrossingSignal = factor(CrossingSignal, levels=c(0, -1, 1)))
```

```{r}
# Focusing on the Age characteristic
df.age_profiles = profiles %>%
  filter(ScenarioType==lev[1] & profiles$ScenarioTypeStrict==lev[1]) %>% 
  mutate(AttributeLevel = factor(AttributeLevel, levels=AttLevels[(4*2):(4*2+1)]),
         Barrier = factor(Barrier, levels=c(0, 1)),
         Intervene = factor(Intervention, levels=c(0, 1))) %>% 
  select(-Intervention) %>% 
  sample_n(10000)
```

```{r message=F}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = b),
  prior(exponential(1), class = sd)
)

model_brms <- brm(
  bf(Saved ~ as.factor(Intervene) + Barrier + CrossingSignal + AttributeLevel + (1 | ResponseID)),
  data = df.age_profiles,
  family = bernoulli(link = "logit"),
  prior = priors,
  chains = 4, cores = 4, iter = 2000, seed = 1234
)
```

```{r}
# Code referenced from: https://www.andrewheiss.com/blog/2023/07/25/conjoint-bayesian-frequentist-guide/#amces-1
posterior_mfx <- model_brms %>% 
  avg_slopes(newdata = "mean", allow_new_levels = TRUE) %>% 
  posterior_draws() 

posterior_mfx_nested <- posterior_mfx %>% 
  separate_wider_delim(
    contrast,
    delim = " - ", 
    names = c("variable_level", "reference_level")
  ) %>% 
  group_by(term, variable_level) %>% 
  nest()

model_variable_levels <- tibble(
  variable = c("AttributeLevel", "Barrier", "CrossingSignal", "Intervene"),
  levels = list(factor(c("Old", "Young"), levels=c("Old", "Young")),
             factor(c(0, 1), levels=c(0, 1)),
             factor(c(-1, 0, 1), levels=c(0, -1, 1)),
             factor(c(0, 1), levels=c(0, 1))
  )) %>%
  unnest(levels)

# Combine full dataset of factor levels with model results
plot_data_bayes <- model_variable_levels %>%
  left_join(
    posterior_mfx_nested,
    by = join_by(variable == term, levels == variable_level)
  ) %>%
  mutate(data = map_if(data, is.null, ~ tibble(draw = 0, estimate = 0))) %>% 
  unnest(data)

label_pp <- label_number(accuracy = 1, scale = 100, 
                         suffix = " pp.", style_negative = "minus")

ggplot(plot_data_bayes, aes(x = draw, y = levels, fill = variable)) +
  geom_vline(xintercept = 0) +
  stat_halfeye(normalize = "groups") +
  guides(fill = "none") +
  facet_col(facets = "variable", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_pp) +
  labs(
    x = "Percentage point change in probability of profile selection",
    y = NULL,
    title = "Posterior AMCEs"
  )
```

## Analysising the data using sum within slices & breakdown

A common data analysis technique is to slice the data by a particular subset of columns and values within those columns then apply SUM (or other metrics) to some measure column.

Which should then theoretically allow us to replicate all the above results without use of lm modeling by simply taking a ratio of `num_saved` / `total` for particular tuples of (Intervene, Barrier, CrossingSignal, AttributeLevel).

Furthermore, to analyze Intervene, Barrier, or CrossingSignal (invertention, passenger vs pedestrain, legality attributes), we can SUM again over different ScenarioTypes and AttributeLevels.

The key is to first precompute the sum of saved profiles and total number of profiles with different categorical levels of attributes.

```{r}
# Precompute categorical levels
df.agg_profiles = profiles %>%
  filter(ScenarioType == ScenarioTypeStrict) %>%
  group_by(Intervention, Barrier, CrossingSignal, PedPed, ScenarioType, AttributeLevel) %>%
  summarize(num_saved = sum(ifelse(Saved == 1, 1, 0)),
            total = n())
```

```{r}
# Attempt to recover age-component causal effect
df.agg_profiles %>%
  filter(ScenarioType == "Age") %>%
  group_by(AttributeLevel) %>% 
  summarize(porportion = sum(num_saved)/sum(total))
```

As a sanity check, the above two proportions sum to 1 (as expected).

This gives a difference of 0.5312954 when setting Old to be the baseline. Which is pretty accurate! Seeing as how we haven't applied weighting to this and the original method returned 0.5097648.

To apply weighting to this, we would multiply the `num_saved` count by some inverse proportion of the total sum of other attributes. 

```{r}
df.agg_profiles = df.agg_profiles %>%
  left_join(df.agg_profiles %>%
      group_by(Intervention, Barrier, CrossingSignal) %>% 
      summarize(wsamples = sum(total)) %>%
      ungroup() %>% 
      mutate(wtotal = sum(wsamples)) %>%
      mutate(wproportion = wsamples/wtotal) %>%
      mutate(weight = 1/wproportion)) %>%
  select(-wsamples,-wtotal,-wproportion)
```

```{r}
# Try again weighted age-component causal effect
df.agg_profiles %>%
  filter(ScenarioType == "Age") %>%
  group_by(AttributeLevel) %>%
  mutate(weight_sum = sum(weight)) %>%
  summarize(porportion = sum(num_saved*(weight/weight_sum))/sum(total*(weight/weight_sum)))
```

Now the difference is 0.5215997.

The weight has slightly moderated the difference between the two levels, however, since I'm using the a posteriori weights (experimental distribution weights) instead of the a priori ones (theoretical ones), there is still some slight differences. Let's try to fix that.

```{r}
df.agg_profiles['weight'] = calcWeightsTheoretical(df.agg_profiles)

# Try again weighted age-component causal effect
df.agg_profiles %>%
  filter(ScenarioType == "Age") %>%
  group_by(AttributeLevel) %>%
  mutate(weight_sum = sum(weight)) %>%
  summarize(porportion = sum(num_saved*(weight/weight_sum))/sum(total*(weight/weight_sum)))
```

This provides us a difference of 0.5097647 \o/ which exactly recovers the original study's computed average marginal causal effect without using lm but also without a measure of clustered errors.
